---
title: "HW3"
subtitle: "Visualizing FEMA NRI x ACS Data"
description: "Assigned Wed 02/04/2026 | Due Wed 02/11/2026"
author: "Austin Martinez"
date: 02/10/2026
format: html
---

## How does climate hazard risk exposure vary across racial / ethnic groups in California?

```{r}
# Import libraries
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

### Data Cleaning

```{r}
# #....Step 1a: see all available ACS variables + descriptions.....
# acs_vars <- tidycensus::load_variables(year = 2023,
#                                        dataset = "acs1")
# 
# #..............Step 1b: import race & ethnicity data.............
# race_ethnicity <- tidycensus::get_acs(
#   geography = "county",
#   survey = "acs1",
#   # NOTE: you may not end up using all these variables
#   variables = c("B01003_001", "B02001_002", "B02001_003", 
#                 "B02001_004", "B02001_005", "B02001_006",
#                 "B02001_007", "B02001_008", "B03002_012",
#                 "B03002_002"),
#   state = "CA", 
#   year = 2023) |>
#   # join variable descriptions (so we know what's what!)
#   dplyr::left_join(acs_vars, by = dplyr::join_by(variable == name)) 
# 
# #.................Step 2: write ACS data to file.................
# readr::write_csv(race_ethnicity, here::here("data", "ACS-1yr-2023-county-race-ethnicity.csv"))

#..................Step 3: read in your CSV file.................
race_ethnicity <- read_csv(here::here("data", "ACS-1yr-2023-county-race-ethnicity.csv"))|> 
    janitor::clean_names()

risk_index <- read_csv(here::here("data", "National_Risk_Index_Counties_807384124455672111.csv"))|> 
    janitor::clean_names()
```

### Data Preporation

```{r}
# Prepare risk_index for join

# Filter for California
CA_risk_index <- risk_index |>
  filter(`state_name` == "California")
```

```{r}
# Prepare race_ethnicity for join

# Remove " County, California$" from every row in name column
race_ethnicity$name <- gsub(" County, California$", "", race_ethnicity$name)

# Remove "!!" from every label column
race_ethnicity$label <- gsub("!!", "", race_ethnicity$label)

# Rename name column to county_name
race_ethnicity  <- race_ethnicity |>
  rename(county_name = name )
```

```{r}
# Left join race_ethnicity to CA_risk_index by county_name
risk_ethnicity_joined <- CA_risk_index |>
  left_join(race_ethnicity, by = "county_name", relationship = "one-to-many") |>
  # Select necessary columns
  select(state_name, county_name, population_2020, national_risk_index_rating_composite, national_risk_index_score_composite, estimate, label, concept)
```

```{r}
# filter our rows with "EstimateTotal" and remove "EstimateTotal:" from all rows in label column
clean_risk_ethnicity_joined<- risk_ethnicity_joined |>
  filter(label != "EstimateTotal") |>
  mutate(label = gsub("EstimateTotal:", "", label))
```

```{r}
# Group by label and calculate percentage of population in each ethnic group in counties with a "very high" risk index
risk_by_race <- clean_risk_ethnicity_joined |>
  group_by(label) |>
  summarize(
    total_pop = sum(estimate, na.rm = TRUE),
    very_high_pop = sum(estimate[national_risk_index_rating_composite == "Very High"], na.rm = TRUE),
    pct_very_high = 100 * very_high_pop / sum(estimate, na.rm = TRUE)
  )

# Rename labels
risk_by_race <- risk_by_race |>
  mutate(label = case_when(
    label == "American Indian and Alaska Native alone" ~ "American Indian and Alaska Native",
    label == "Asian alone" ~ "Asian",
    label == "Black or African American alone" ~ "Black or African American",
    label == "Hispanic or Latino:" ~ "Hispanic or Latino",
    label == "Native Hawaiian and Other Pacific Islander alone" ~ "Native Hawaiian and Other Pacific Islander",
    label == "Not Hispanic or Latino:" ~ "Not Hispanic or Latino",
    label == "Some Other Race alone" ~ "Some Other Race", 
    label == "White alone" ~ "White", 
    label == "Two or More Races:" ~ "Two or More Races",
    TRUE ~ label
    ))
```

### Data Visualization

```{r fig.height=6, fig.width=13}
#| fig-alt: "Bar plot of the percentage of different ethnic groups living in counties with 'Very High' FEMA National Risk Index scores for all U.S. states, where Asians have the highest percentage of people living in 'Very High' Risk Counties."

risk_by_race |>
  ggplot(aes(x = reorder(label, pct_very_high), y = pct_very_high)) +
  geom_col(fill = "firebrick") +
  labs(
    title = "Percent of Population Living in 'Very High' Risk Counties",
    subtitle = "Asians have the highest percentage of people living in 'Very High' Risk Counties",
    y = "Percent of Population in Very High Risk Counties (%)",
    caption = "Data: FEMA National Risk Index (2023 Release)"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(axis.title.y = element_blank()) +
  geom_text(aes(label = round(pct_very_high, 1), hjust = -0.4))
```

## V. Answer some questions

1.  What are your variables of interest and what kinds of data (e.g. numeric, categorical, ordered, etc.) are they (a bullet point list is fine)?

variables in data set: - (label) – categorical

-   (county_name) – categorical

-   (estimate) – numeric

-   (national_risk_index_rating_composite) – categorical

variables I made: - (very_high_pop) – numeric

-   (pct_very_high) – numeric

2.  How did you decide which type of graphic form was best suited for answering the question? What alternative graphic forms could you have used instead? Why did you settle on this particular graphic form?

-   I decided to choose a bar graph because it clearly shows the difference in populations between ethnic groups living in "Very High" risk counties. I alternatively could have chosen a lollipop graph, but I don't think its necessary because there aren't that many groups. This type of a graph is more familiar to most people, so its more comprehensible.

3.  Summarize your main finding in no more than two sentences.

-   Asians have the highest percentage of people living in 'Very High' Risk Counties. Pacific islanders/Native Hawaiians and White people have the least percentage of people living in 'Very High' Risk Counties.

4.  What modifications did you make to this visualization to make it more easily readable?

-   I made my plot wider and taller, made my bars dark red, put ethnic groups on the y axis, and added percentage values to the end of each bar.

5.  Is there anything you wanted to implement, but didn’t know how? If so, please describe.

-   I want to make the text on y axis the same size as y axis.
